quollr: An R Package for Visualizing $2-\\text{D}$ Models from Non-linear Dimension Reductions in High Dimensional Space

### 2025-05-08 paper-quollr.Rmd
Added updated version

nldr_scaled_obj <- gen_scaled_data(
  data = scurve_umap1)
umap_scaled1 <- nldr_scaled_obj$scaled_nldr

@@

nldr_scaled_obj2 <- gen_scaled_data(
  data = scurve_umap2)
umap_scaled2 <- nldr_scaled_obj2$scaled_nldr

@@

nldr_scaled_obj3 <- gen_scaled_data(
  data = scurve_umap3)
umap_scaled3 <- nldr_scaled_obj3$scaled_nldr

@@

nldr_scaled_obj4 <- gen_scaled_data(
  data = scurve_umap4)
umap_scaled4 <- nldr_scaled_obj4$scaled_nldr

@@

nldr_scaled_obj5 <- gen_scaled_data(
  data = scurve_umap5)
umap_scaled5 <- nldr_scaled_obj5$scaled_nldr

@@

nldr_scaled_obj6 <- gen_scaled_data(
  data = scurve_umap6)
umap_scaled6 <- nldr_scaled_obj6$scaled_nldr

@@

`scurve` is a $7\text{-}D$ simulated dataset. It is constructed by simulating $5000$ observations from $\theta \sim U(-3\pi/2, 3\pi/2)$, $X_1 = \sin(\theta)$, $X_2 \sim U(0, 2)$ (adding thickness to the S), $X_3 = \text{sign}(\theta) \times (\cos(\theta) - 1)$. The remaining variables $X_4, X_5, X_6, X_7$ are all uniform error, with small variance. `scurve_umap` is the UMAP $2\text{-}D$ embedding for `scurve` data with `n_neighbors` is $15$ and `min_dist` is $0.1$. Each data set contains a unique ID column that maps `scurve` and `scurve_umap`.

@@

```{r, echo=TRUE}
calc_bins_y(
  bin1 = 12, 
  r2 = r2,
  q = 0.1
  )
```

@@

```{r, echo=TRUE}
hex_binning(
  data = s_curve_noise_umap_scaled, 
  bin1 = 12, 
  r2 = r2,
  q = 0.1
  )
```

@@

```{r, echo=TRUE, eval=FALSE}
find_low_dens_hex(
  df_bin_centroids_all = df_bin_centroids, 
  bin1 = 12, 
  df_bin_centroids_low = df_bin_centroids_low
  )

@@

```{r, echo=TRUE, eval=FALSE}
find_lg_benchmark(
  distance_edges = distance_df, 
  distance_col = "distance"
  )
```

@@

The final step involves lifting the fitted $2\text{-}D$ model into $p\text{-}D$ by computing the $p\text{-}D$ mean of data points within each bin to represent bin centroids. This transformation is performed using the `avg_highd_data()` function, which takes $p\text{-}D$ data and their corresponding hexagonal bin IDs as inputs.

```{r, echo=TRUE, eval=FALSE}
umap_data_with_hb_id <- hb_obj$data_hb_id

df_all <- bind_cols(
  s_curve_noise_training |> dplyr::select(-ID), 
  umap_data_with_hb_id
  )

df_bin <- avg_highd_data(
  data = df_all
  )
```

@@

```{r, echo=TRUE, eval=FALSE}
predict_emb(
  highd_data = s_curve_noise_training, 
  model_2d = df_bin_centroids, 
  model_highd = df_bin
  )
```

@@

```{r, echo=TRUE, eval=FALSE}
glance(
  highd_data = s_curve_noise_training, 
  model_2d = df_bin_centroids, 
  model_highd = df_bin
  )
```

@@

```{r, echo=TRUE, eval=FALSE}
augment(
  highd_data = s_curve_noise_training, 
  model_2d = df_bin_centroids, 
  model_highd = df_bin
  )
```

@@

```{r, echo=TRUE, eval=FALSE}
vis_lg_mesh(
  distance_edges = distance_df, 
  benchmark_value = 0.75, 
  tr_coord_df = tr_from_to_df, 
  distance_col = "distance"
  )
```

```{r, echo=TRUE, eval=FALSE}
vis_rmlg_mesh(
  distance_edges = distance_df, 
  benchmark_value = 0.75, 
  tr_coord_df = tr_from_to_df, 
  distance_col = "distance"
  )
```

@@

```{r, echo=TRUE, eval=FALSE}
show_langevitour(
  df = df_all, 
  df_b = df_bin, 
  df_b_with_center_data = df_bin_centroids, 
  benchmark_value = 0.75, 
  distance = distance_df, 
  distance_col = "distance", 
  use_default_benchmark_val = FALSE, 
  col_start = "x"
  )
```

>>>>

nldr_scaled_obj <- gen_scaled_data(
  nldr_data = scurve_umap)
umap_scaled1 <- nldr_scaled_obj$scaled_nldr

@@

nldr_scaled_obj2 <- gen_scaled_data(
  nldr_data = scurve_umap2)
umap_scaled2 <- nldr_scaled_obj2$scaled_nldr

@@

nldr_scaled_obj3 <- gen_scaled_data(
  nldr_data = scurve_umap3)
umap_scaled3 <- nldr_scaled_obj3$scaled_nldr

@@

nldr_scaled_obj4 <- gen_scaled_data(
  nldr_data = scurve_umap4)
umap_scaled4 <- nldr_scaled_obj4$scaled_nldr

@@

nldr_scaled_obj5 <- gen_scaled_data(
  nldr_data = scurve_umap5)
umap_scaled5 <- nldr_scaled_obj5$scaled_nldr

@@

nldr_scaled_obj6 <- gen_scaled_data(
  nldr_data = scurve_umap6)
umap_scaled6 <- nldr_scaled_obj6$scaled_nldr

@@

`scurve` is a $7\text{-}D$ simulated dataset. It is constructed by simulating $5000$ observations from $\theta \sim U(-3\pi/2, 3\pi/2)$, $X_1 = \sin(\theta)$, $X_2 \sim U(0, 2)$ (adding thickness to the S), $X_3 = \text{sign}(\theta) \times (\cos(\theta) - 1)$. The remaining variables $X_4, X_5, X_6, X_7$ are all uniform error, with small variance. `scurve_umap` is the UMAP $2\text{-}D$ embedding for `scurve` data with `n_neighbors` is $46$ and `min_dist` is $0.9$. Each data set contains a unique ID column that maps `scurve` and `scurve_umap`.

@@

```{r, echo=TRUE}
calc_bins_y(
  nldr_obj = scurve_nldr_obj, 
  bin1 = 4, 
  q = 0.1)
```

@@

```{r, echo=TRUE}
hex_binning(
  nldr_obj = scurve_nldr_obj, 
  bin1 = 4, 
  q = 0.1)
```

@@

@@

```{r, echo=TRUE, eval=FALSE}
find_low_dens_hex(
  centroids_data, 
  bin1 = 15, 
  benchmark_mean_dens = 0.05
)

@@

The final step involves lifting the fitted $2\text{-}D$ model into $p\text{-}D$ by computing the $p\text{-}D$ mean of data points within each hexgaonal bin to represent bin centroids. This transformation is performed using the `avg_highd_data()` function, which takes $p\text{-}D$ data (`highd_data`) and embedding data with their corresponding hexagonal bin IDs as inputs (`scaled_nldr_hexid`).

```{r, echo=TRUE, eval=FALSE}
df_bin <- avg_highd_data(
  highd_data = scurve, 
  scaled_nldr_hexid
)
```

@@

```{r, echo=TRUE, eval=FALSE}
predict_emb(
  highd_data = scurve, 
  model_2d = df_bin_centroids, 
  model_highd = df_bin
  )
```

@@

```{r, echo=TRUE, eval=FALSE}
glance(
  highd_data = scurve, 
  model_2d = df_bin_centroids, 
  model_highd = df_bin
  )
```

@@

```{r, echo=TRUE, eval=FALSE}
augment(
  highd_data = scurve, 
  model_2d = df_bin_centroids, 
  model_highd = df_bin
  )
```

@@

```{r, echo=TRUE, eval=FALSE}
vis_mesh(
  distance_edges = distance_df, 
  benchmark_value = 0.75, 
  tr_coord_df = tr_from_to_df, 
  distance_col = "distance"
  )
```

@@

```{r, echo=TRUE, eval=FALSE}
df_exe <- comb_data_mode(
  highd_data = scurve, 
  model_highd = scurve_model_obj$model_highd, 
  model_2d = scurve_model_obj$model_2d
  )
```

```{r, echo=TRUE, eval=FALSE}
show_langevitour(
  point_df, 
  edge_df
  )
```
